---
title: "Lab 2"
author: "Mason Sharum"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1

```{r}
getwd()
```

## Task 2

```{r}
epa = read.csv("EPAGAS-1.csv")
head(epa)
```

## Task 3

### Question 1

```{r}
mpg = epa$MPG
z=(mpg - mean(mpg))/sd(mpg)
round(mean(z), 9)
sd(z) * sd(z)
```

### Question 2

```{r}
epa[abs(z) >= 2 & abs(z) <= 3,]
```

### Question 3

```{r}
epa[abs(z) > 3,]
```

### Question 4

```{r}
library(lattice)
mycol = ifelse(abs(z) > 3, "Red",
          ifelse(abs(z) >= 2 & abs(z) <= 3, "Blue", "Black"))
dotplot(mpg, col = mycol)
```

## Task 4

### Question 1

```{r}
boxplot(mpg, main = "MPG Variable Boxplot", notch = TRUE, horizontal = TRUE, xlab = "Miles Per Gallon", col = "Black")
```

### Question 2

Using Chebyshev's theorem there should be at least 3/4 of the data within 2 standard deviations of the mean.

### Question 3

```{r}
length(mpg[abs(z)<2])/length(mpg)
```

### Question 4

Chebyshev does agree with the data since 96/100 is more than 3/4.

### Question 5

The empirical rule requires there to be at least  95% of the data within 2 standard deviations of the mean.

### Question 6

```{r}
length(mpg[abs(z)<2])/length(mpg)
```

It corresponds well since there is 96% of the data within 2 standard deviations.

### Question 7

```{r}
hist(mpg)
```

Yes, the Empirical rule is valid since mpg is unimodal and symmetric about the mode.