---
title: "Lab 1"
author: "Mason Sharum"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1

```{r}
getwd()
```

## Task 2

```{r}
ddt = read.csv("DDT.csv")
head(ddt)
```

## Task 3

```{r}
summary(ddt)
```
### Question 1

The qualitative variables in "ddt" are River and Species.

### Question 2

The quantitative variables in "ddt" are Mile, Length, Weight, and DDT.

### Question 3

```{r}
table(ddt$SPECIES)
```

There are 3 SPECIES in the ddt data set.

### Question 4

```{r}
with(ddt, ddt[WEIGHT>800 & SPECIES=="LMBASS",])
```

### Question 5

```{r}
with(ddt, ddt[RIVER=="SCM" & DDT>4.0,])
```

## Task 4

### Question 1

```{r}
mean(ddt$LENGTH)
```
The mean length of fish is 42.8125.

### Question 2

```{r}
sd(ddt$WEIGHT)
```
The standard deviation of the weight of fish is 376.5461.

### Question 3

```{r}
with(ddt, plot(LENGTH ~ WEIGHT))
```

Answer: No, the given plot is not the same as the length vs. weight from the ddt data frame.

### Question 4

```{r}
v = 1:20
table(v/20)
```
The last value of v/20 is 1.

## Task 5

### Question 1

```{r}
rivers = with(ddt, table(RIVER))
rivers
```

### Question 2

```{r}
barplot(rivers, col=1:4)
```

### Question 3

```{r}
SPxRIV = with(ddt, table(RIVER, SPECIES))
SPxRIV
```

### Question 4

```{r}
barplot(SPxRIV)
```

## Task 6

### Question 1

```{r}
species = with(ddt, table(SPECIES))
pie(species)
```

### Question 2

```{r}
pie(rivers)
```

## Task 7

### Question 1

```{r}
vddt = with(ddt, table(DDT))
boxplot(vddt, ylab = "DDT")
```

### Question 2

```{r}
weight = with(ddt, table(WEIGHT))
boxplot(weight, ylab = "Weight")
```

### Question 3

```{r}
length = with(ddt, table(LENGTH))
boxplot(length, ylab = "Length")
```

## Task 8

### Question 1

```{r}
with(ddt, coplot(LENGTH ~ WEIGHT | RIVER))
```

### Question 2

```{r}
with(ddt, coplot(DDT ~ WEIGHT | SPECIES))
```

## Task 9

### Question 1

```{r}
library(ggplot2)
ggplot(ddt, aes(x = SPECIES, y = WEIGHT, fill = RIVER)) + geom_boxplot() + ggtitle("Mason Sharum")
```

### Question 2

```{r}
ggplot(ddt, aes(x = RIVER, y = LENGTH, fill = SPECIES)) + geom_violin() + ggtitle("Mason Sharum")
```

### Question 3

```{r}
ggplot(ddt, aes(x = WEIGHT, y = LENGTH, color = SPECIES)) + geom_point() + ggtitle("Mason Sharum")
```