---
title: "Assignment 1"
author: "Mason Sharum"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I completed 15/15 questions in this assignment.

# Question 1

Below is the table of grading weights for the course.

| Group            | Amount | Total Weight |
| ---------------  | -----: | -----------: |
| Assignments      | 4      | 15%          |
| Laboratories     | 16     | 10%          |
| Projects         | 2      | 10%          |
| In Class Quizzes | TBD    | 10%          |
| Chapter Quizzes  | TBD    | 5%           |
| Mid-Term Exam    | 1      | 20%          |
| Final Exam       | 1      | 30%          |

The grading scale is as follows: A (100-90) B(89-80) C(79-60) D(59-50) F(<50), and there will be no curves.


# Question 2

```{r}
ddt <- read.csv("DDT.csv")
head(ddt)

ddt$MILE = factor(ddt$MILE)
```

## (a)

```{r}
with(ddt, coplot(LENGTH ~ WEIGHT | RIVER * SPECIES, ddt, col = MILE))
```

## (b)

The length and weight of catfish found in the FCM, LCM, and SCM rivers increase together in a positive correlation.

## (c)

Line A is a function call that creates a factor object of the MILE variable, meaning it makes a vector of the measurements from MILE after sorting them into categories.

## (d)

Line B is a function call that returns the number of unique values in the previously made vector. For instance if the vector was [1, 1, 2, 3, 3], it would return 3 since there are three unique numbers.

## (e)

The top six plots are empty because there are no fish that are either LMBASS or SMBUFFALO that were caught in the FCM, LCM, or SCM rivers.

## (f)

```{r}
cat = subset(ddt, RIVER == "FCM" & SPECIES == "CCATFISH",)
mean(cat$DDT)
```

# Question 3

## (a)

Length of the maximum span (feet) is a quantitative variable.

## (b)

Number of vehicle lanes is a quantitative variable.

## (c)

TOll bridge (yes or no) is a qualitative variable.

## (d)

Average daily traffic is a quantitative variable.

## (e)

Condition of deck (good, fair, poor) is a qualitative variable.

## (f)

Bypass or detour length (miles) is a quantitative variable.

## (g)

Route type (interstate, U.S., state, county, or city) is a qualitative variable.

# Question 4

## (a)

The four random sampling designs are as follows: simple random sampling, stratified random sampling, cluster sampling, and systematic sampling.

## (b)

Simple random sampling happens when n experimental units are selected from the population in such a way that every different sample of size n has an equal chance of selection.

Stratified random sampling happens when the population is separated into different strata based off of certain characteristics of the experimental units. These strata are then randomly sampled from.

Cluster sampling happens when a group(or cluster) is selected from within the population and then data is taken about every unit in the cluster.

Systematic sampling happens when every kth experimental unit is selected from a list of all experimental units.

# Question 5

```{r}
mtbe <- read.csv("MTBE.csv")
head(mtbe)

ind = sample(1:223, 5, replace = FALSE)
mtbe[ind,]
```

## (a)

### (i)

```{r}
mtbeo = na.omit(mtbe)
```

### (ii)

```{r}
depth = mtbeo[mtbeo$Aquifier == "Bedrock",]$Depth
sd(depth)
```

# Question 6

```{r}
eq <- read.csv("EARTHQUAKE.csv")
head(eq)

ind = sample(1:2929, 30, replace = FALSE)
eq[ind,]
```

## (a)

### (i)

```{r}
plot(ts(eq$MAG))

median(eq$MAGNITUDE)
```

# Question 7

## (a)

The data was collected with a stratified sample.

## (b)

The population is all of the fish in the Tennessee River.

## (c)

The qualitative variables are location and species. Location was divided into river and mile.

# Question 8

## (a)

A bar graph is used to describe the data.

## (b)

The variable measured for each of the 106 robot designs was type of robotic limbs.

## (c)

The current most used design of robot is legs only.

## (d)

```{r}
limb_freq = c(15, 8, 63, 20)
limb_labels = c("None", "Both", "Legs", "Wheels")
rel_freq = as.data.frame(matrix(data = limb_freq/sum(limb_freq), nrow = 4, ncol = 1), row.names = limb_labels)
rel_freq
```

## (e)

```{r}
source("pareto.r")
l = rep(limb_labels, limb_freq)
pareto(l)
```

# Question 9

## (a)

```{r}
nums = c(32, 6, 12)
labs = c("Windows", "Explorer", "Office")
pie(nums, labels = labs, main = "Microsoft Products with Security Issues")
```

## (b)

```{r}
MSFreq = c(6, 8, 22, 3, 11)
MSLabel = c("DoS", "Info Leak", "RCE", "Spoofing", "Elevation")
MS = rep(MSLabel, MSFreq)
pareto(MS)
```

Based on this diagram, I would say they should focus on fixing the Remote Code Execution problem, as that seems to be the most commonly used exploit.

# Question 10

```{r}
swd = read.csv("SWDEFECTS.csv")
head(swd)

library(plotrix)
tab = table(swd$defect)
rtab = tab/sum(tab)
round(rtab, 2)
pie3D(rtab, labels = list("OK", "Defective"), main = "Pie Plot of SWD")
```

Based off of the pie chart it is easy to see that there is a high likelihood of defective software code.

# Question 11



# Question 12

```{r}
pipe <- read.csv("ROUGHPIPE.csv")
avg = mean(pipe$R)
stddev = sd(pipe$R)
c(avg - (2*stddev), avg + (2*stddev))
```

This is the range of two standard deviations away from the mean in each direction, which according to the empirical rule will contain approximately 95% of the data since the shape of the data is normal.

# Question 13

## (a)

```{r}
ants = read.csv("GOBIANTS.csv")
head(ants)
source("getmode.R")
mean(ants$AntSpecies)
median(ants$AntSpecies)
getmode(ants$AntSpecies)
```
The mean is the average number of species found. The median is the middle number of species found. The mode is the most common number of species found.

## (b)

I would use the median as the mean is much higher than the median meaning the mean is probably skewed. The mode is 
just the most common number which could be the largest number appearing twice and hundreds of smaller numbers only appearing once. The median is always in the middle of the data, and in this case seems to be the best measure for this.

## (c)



# Question 14



# Question 15

```{r}
library(ggplot2)
ggplot(ddt, aes(x = RIVER, y = LENGTH)) + geom_boxplot(aes(fill = SPECIES)) + ggtitle("Mason Sharum")
```






















