---
title: "Assignment 1"
author: "Mason Sharum"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I completed 15/15 questions in this assignment.

# Question 1

Below is the table of grading weights for the course.

| Group            | Amount | Total Weight |
| ---------------  | -----: | -----------: |
| Assignments      | 4      | 15%          |
| Laboratories     | 16     | 10%          |
| Projects         | 2      | 10%          |
| In Class Quizzes | TBD    | 10%          |
| Chapter Quizzes  | TBD    | 5%           |
| Mid-Term Exam    | 1      | 20%          |
| Final Exam       | 1      | 30%          |

The grading scale is as follows: A (100-90) B(89-80) C(79-60) D(59-50) F(<50), and there will be no curves.


# Question 2

```{r}
ddt <- read.csv("DDT.csv")
head(ddt)

ddt$MILE = factor(ddt$MILE)
```

## (a)

```{r}
with(ddt, coplot(LENGTH ~ WEIGHT | RIVER * SPECIES, ddt, col = MILE))
```

## (b)

The length and weight of catfish found in the FCM, LCM, and SCM rivers increase together in a positive correlation.

## (c)

Line A is a function call that creates a factor object of the MILE variable, meaning it makes a vector of the measurements from MILE after sorting them into categories.

## (d)

Line B is a function call that returns the number of unique values in the previously made vector. For instance if the vector was [1, 1, 2, 3, 3], it would return 3 since there are three unique numbers.

## (e)

The top six plots are empty because there are no fish that are either LMBASS or SMBUFFALO that were caught in the FCM, LCM, or SCM rivers.

## (f)

```{r}
cat = subset(ddt, RIVER == "FCM" & SPECIES == "CCATFISH",)
mean(cat$DDT)
```

# Question 3

## (a)

Length of the maximum span (feet) is a quantitative variable.

## (b)

Number of vehicle lanes is a quantitative variable.

## (c)

TOll bridge (yes or no) is a qualitative variable.

## (d)

Average daily traffic is a quantitative variable.

## (e)

Condition of deck (good, fair, poor) is a qualitative variable.

## (f)

Bypass or detour length (miles) is a quantitative variable.

## (g)

Route type (interstate, U.S., state, county, or city) is a qualitative variable.

# Question 4

## (a)

The four random sampling designs are as follows: simple random sampling, stratified random sampling, cluster sampling, and systematic sampling.

## (b)

Simple random sampling happens when n experimental units are selected from the population in such a way that every different sample of size n has an equal chance of selection.

Stratified random sampling happens when the population is separated into different strata based off of certain characteristics of the experimental units. These strata are then randomly sampled from.

Cluster sampling happens when a group(or cluster) is selected from within the population and then data is taken about every unit in the cluster.

Systematic sampling happens when every kth experimental unit is selected from a list of all experimental units.

# Question 5

```{r}
mtbe <- read.csv("MTBE.csv")
head(mtbe)

ind = sample(1:223, 5, replace = FALSE)
mtbe[ind,]
```

## (a)

### (i)

```{r}
mtbeo = na.omit(mtbe)
```

### (ii)

```{r}
depth = mtbeo[mtbeo$Aquifier == "Bedrock",]$Depth
sd(depth)
```

# Question 6

```{r}
eq <- read.csv("EARTHQUAKE.csv")
head(eq)

ind = sample(1:2929, 30, replace = FALSE)
eq[ind,]
```

## (a)

### (i)

```{r}
plot(ts(eq$MAG))

median(eq$MAGNITUDE)
```

# Question 7

## (a)

The data was collected with a stratified sample.

## (b)

The population is all of the fish in the Tennessee River.

## (c)

The qualitative variables are location and species. Location was divided into river and mile.

# Question 8

## (a)

A bar graph is used to describe the data.

## (b)

The variable measured for each of the 106 robot designs was type of robotic limbs.

## (c)

The current most used design of robot is legs only.

## (d)

```{r}
limb_freq = c(15, 8, 63, 20)
limb_labels = c("None", "Both", "Legs", "Wheels")
rel_freq = as.data.frame(matrix(data = limb_freq/sum(limb_freq), nrow = 4, ncol = 1), row.names = limb_labels)
rel_freq
```

## (e)

```{r}
source("pareto.r")
l = rep(limb_labels, limb_freq)
pareto(l)
```

# Question 9

## (a)

```{r}
nums = c(32, 6, 12)
labs = c("Windows", "Explorer", "Office")
pie(nums, labels = labs, main = "Microsoft Products with Security Issues")
```

## (b)

```{r}

```























